Optimizing the power of adaptive seamless designs using Bayesian Optimization

### 2020-07-24 main.tex
Updates from Overleaf

% max 250 words
\begin{abstract}
Planning the patient group sizes and test methids for clinical trials in such a way that the test power is as high as possible for given effect sizes under the null hypothesis and for a fixed number of treatments is a crucial task.
From a wide set of possible set ups we want to chose the most promising plan in short time to drive fast decisions.
While it is possible to simulate the power for all various set ups, the space of possible set ups soon becomes too big.
Either you need big computational resources or you have to wait for an exhausive exploration of the possible set ups.
In this paper we show that you can successfully apply Model-based Optimization to find a clinical trial set-up that yields an high power from a wide set of possible set-ups.
We optimize the power of adaptive seamless designs for different sets of treatment effect sizes and compare the result to an exhausitve search to show that our optimization approach finds a promising set-up in a fraction of the time.
\end{abstract}

@@

Clinical research and in particular drug development is typically structured into phases, e.g. phases I to IV in drug development; similar approaches for e.g. complex interventions (MRC framework). 
The delepment phases include elements of learning and comfirmation. 
For instance, Sheiner (1997) described the drug development process as two learning and confirming cycles. 
First learning / confirming cycle (Phase I-IIa): Learning about tolerated dose (Phase I). 
Then confirming of efficacy of selected dose in selected group of patients (Phase IIa). 
Second learning / confirming cycle (Phase IIb-III): Learning about optimal use in respresentative patients (Phase IIb); Then confirming of acceptable benefit / risk ratio (Phase III). 
Traditionally, separate studies for learning and confirming.

@@

\section{Motivating Example}
% TODO: Jacob 
COPD trial
% Friede et al (2020) Biom J, Section 5.1 Clinical trial in COPD with treatment selection; other references: Barnes et al. (2011) and Donohue et al. (2010); Cuffe, Lawrence, Stone,&Vandemeulebroecke, 2014.

@@

% \begin{thebibliography}{10}
% \bibitem[Bauer and Bauer(1994)Bauer, P. and Bauer, M.M.]{bib1}Bauer, P. and  Bauer, M. M. (1994). Testing equivalence simultaneously for location and  dispersion of two normally distributed populations.  \textit{Biometrical  Journal} \textbf{36}, 643--660.
% \bibitem[Farrington, C. P. and Andrews, N. (2003)]{bib2}Farrington, C.P. and Andrews, N. (2003). Outbreak detection:
% Application to infectious disease surveillance. In: Monitoring the Health of Populations (eds. R. Brookmeyer and D. F. Stroup), Oxford University Press, Oxford,\break 203--231.
% \bibitem[Rencher(1998)Rencher, A.C.]{bib3}Rencher, A. C. (1998).  \textit{Multivariate Statistical Inference and Applications}. Wiley, New  York. 
% \end{thebibliography}

\bibliographystyle{biometrical}
\bibliography{../literature.bib}

>>>>

% max 250 words
\begin{abstract}
Planning the patient group sizes and test methods for clinical trials in such a way that the test power is as high as possible for given effect sizes under the null hypothesis and for a fixed number of treatments is a crucial task.
From a wide set of possible set ups we want to chose the most promising plan in short time to drive fast decisions.
While it is possible to simulate the power for all various set ups, the space of possible set ups soon becomes too big.
Either you need big computational resources or you have to wait for an exhaustive exploration of the possible set ups.
In this paper we show that you can successfully apply Model-based Optimization to find a clinical trial set-up that yields an high power from a wide set of possible set-ups.
We optimize the power of adaptive seamless designs for different sets of treatment effect sizes and compare the result to an exhaustive search to show that our optimization approach finds a promising set-up in a fraction of the time.
\end{abstract}

@@

Clinical research and in particular drug development is typically structured into phases, e.g. phases I to IV in drug development; similar approaches for e.g. complex interventions (MRC framework). 
The delepment phases include elements of learning and confirmation. 
For instance, \cite{sheiner_learning_1997} described the drug development process as two learning and confirming cycles. 
First learning / confirming cycle (Phase I-IIa): Learning about tolerated dose (Phase I). 
Then confirming of efficacy of selected dose in selected group of patients (Phase IIa). 
Second learning / confirming cycle (Phase IIb-III): Learning about optimal use in respresentative patients (Phase IIb); Then confirming of acceptable benefit / risk ratio (Phase III). 
Traditionally, separate studies for learning and confirming.

@@

\section{Motivating Example}
% TODO: Jacob 
COPD trial
% Friede et al (2020) Biom J, Section 5.1 Clinical trial in COPD with treatment selection;
\cite{friede_adaptive_2020}

%other references: Barnes et al. (2011) and Donohue et al. (2010); 
\cite{cuffe_when_2014}

@@

\bibliographystyle{biometrical}
\bibliography{literature.bib}
